---
title: "TDP43 clones analysis"
author: "Matteo Zanovello"
date: "2022-01-10"
output:
---

```{r setup, include = False}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
if (!require("pacman")) {
    install.packages("pacman") }
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager") }
pacman::p_load(tidyverse, data.table, ggplot2, janitor, ggpubr, ggrepel,
               devtools, readxl, DESeq2, readr, edgeR, annotables,
               PCAtools, clusterProfiler, "org.Hs.eg.db", enrichplot, pcaExplorer, topGO)

source(here::here('scripts','feature_count','create_feature_count_table.R'))
source(here::here('scripts','feature_count','make_deseq_dfs.R'))
source(here::here('scripts','feature_count','run_standard_deseq.R'))
source(here::here('scripts','feature_count','make_volcano_plot.R'))
source(here::here('scripts','splicing_volcanoplot.R'))
```

this organizes all the data from the many .bam files and creates one data frame you can work with,
then runs the make_deseq_dfs() function and run_standard_deseq() and eventually make graphs
```{r}
featureCounts <- create_feature_count_table("/Users/matteozanovello/Documents/phd/research_lines/tdp43_clones/data/feature_count") #change here!!
#my_df <- make_deseq_dfs(featureCounts, grep_pattern = "", base_grep = "NT", contrast_grep = "DOX")

grepper <- c("_6_", ####nothing
             "_1_",
             "_10_", ###higher than 1 - not really!!!
             "_17_") ###higher than 10


for (j in grepper) {
my.dds <- run_standard_deseq("/Users/matteozanovello/Documents/phd/research_lines/tdp43_clones/data/feature_count", #change here!!
                             base_grep = "NT",
                             contrast_grep = "DOX",  
                             grep_pattern = j,
                             baseName = "Untreated",
                             contrastName = 'TDP43KD')

bbb <- label_significant(my.dds$results_table, log2FoldCut = 2, log10padj = 10)
plot(bbb)
}
```

splicing volcano
```{r}

input <- c("Control6-TDP43KD6", #36%
           "Control1-TDP43KD1", #20%
           "Control10-TDP43KD10", #40%
           "Control17-TDP43KD17")

for (i in input) {
    input_csv <- paste(i, "csv", sep = ".")
input_splicing <- fread(file.path(here::here(), "data", input_csv))
input_splicing <- separate_rows(input_splicing, c(names(input_splicing[,c(4:9,12,15)])), sep  = c(";"), convert = T)

a <- splicing_dots_tables_function(input_splicing) + annotate("text", x=0.9, y=0.1, label=i)
plot(a)
}
```

This is the vignette for the PCAtools package
https://bioconductor.org/packages/release/bioc/vignettes/PCAtools/inst/doc/PCAtools.html

We're going to put the "ensgene" into a 'rowname' and then do the PCA analysis using meta data
Now I'm going to write this using the pipe operator "%>%". Your "hot key" to type that in is: Ctrl + Shift + M
```{r}
#featureCounts <- featureCounts[,c(1,5:13,17:20)]
colnames(featureCounts) <- c("ensgene", 
                           "DOX_1_1", "DOX_1_2", "DOX_1_3", "DOX_1_4",
                           "DOX_10_1", "DOX_10_2", "DOX_10_3", "DOX_10_4",
                           "DOX_17_1", "DOX_17_2", "DOX_17_3", "DOX_17_4",
                           "DOX_6_1", "DOX_6_2", "DOX_6_3", "DOX_6_4",
                           "NT_1_1", "NT_1_2", "NT_1_3", "NT_1_4",
                           "NT_10_1", "NT_10_2", "NT_10_3", "NT_10_4", 
                           "NT_17_1", "NT_17_2", "NT_17_3", "NT_17_4", 
                           "NT_6_1", "NT_6_2", "NT_6_3", "NT_6_4", 
                           "gene_name")

pca_table <- featureCounts[,1:33] %>%
    column_to_rownames('ensgene')
    
path_to_meta_you_downloaded <-  "/Users/matteozanovello/Documents/phd/research_lines/tdp43_clones/data/metadata.csv" #update this
meta_df <- fread(path_to_meta_you_downloaded)
meta_df <- meta_df %>%
    column_to_rownames('sample') 
meta_df[,1] <- as.factor(meta_df[,1])
meta_df[,2] <- as.factor(meta_df[,2])

annotation <- featureCounts[,c(1,34)] %>%
    column_to_rownames('ensgene')

pcaExplorer(countmatrix = pca_table, coldata = meta_df, annotation = annotation)

```

GO analysis https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-go.html
#see notes above
```{r}
results <- my.dds$results_table
filtered_res <- filter(results, padj < 0.1 & abs(log2FoldChange) > 2)
## feature 1: numeric vector
geneList <- pull(filtered_res,3)
## feature 2: named vector
names(geneList) = as.character(pull(filtered_res,8))
## feature 3: decreasing order
geneList = sort(geneList, decreasing = TRUE)

ggo <- groupGO(gene = names(geneList),
               OrgDb = org.Hs.eg.db,
               keyType = "SYMBOL",
               ont = "MF",
               level = 3)
cnetplot(ggo)
```

```{r}
ego <- enrichGO(gene = names(geneList),
                keyType = "SYMBOL",
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.01,
                qvalueCutoff = 0.05)
goplot(ego)
```

```{r}
pgo <- pairwise_termsim(ego)
emapplot(pgo)
```

```{r}
gse <- gseGO(geneList     = geneList,
              keyType = "SYMBOL",
              OrgDb        = org.Hs.eg.db,
              ont          = "CC",
              minGSSize    = 50,
              maxGSSize    = 200,
              pvalueCutoff = 0.01,
              verbose      = FALSE)
```